Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-22 VMCAI-revised-version/final.tex
effectivity

Surprisingly, the general undecidability statement regarding resilience had neither been known nor proven.
We show that the resilience problem is undecidable for $\omega^2$-WSTS with strong compatibility. 
%
% \alain{pourquoi n'a-t-on pas prouvé la même chose pour la BR et la k-R ?: où est la difficulté ?}
\begin{proposition}\label{indec WSTS}
{\sc Resilience} for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
\end{proposition}
 \alain{peut-on ajouter des hypothèses d'effectivity ? peut-on montrer que l'exemple n'est pas pred effectif ? l'exemple est-il  ideal-effective (oui pour les propriétes 1 et 2 mais pas 3 car ça permettrait de décider l'arrêt ?)? on voit que l'exemple n'est pas completion-$\post$-effective vu le theoreme qui suit}
\begin{proof}
Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
increasing recursive functions from \cite{FMP-ic04} defined as

@@

\alain{énoncer pour les reset-VASS c'est plus fort}
\begin{proposition}\label{srp up down}
{\sc State-resilience},
{\sc Bounded-state-resilience} and
{\sc $k$-state-resilience}
are undecidable for strongly compatible,
$\post$-effective, $\pred$-basis-effective WSTS
when
$\Safe=\mathop{\uparrow} \Safe$.
\end{proposition}

@@

We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
%
Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines.
Indeed, {\sc State-resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
 As seen in 
Proposition~\ref{indec WSTS with dcs},
 the sequences of $M$ reaching a $\Safe$ state from a $\Bad$ one can be simulated in a reset-VASS
by sequences reaching a $\Safe'$ state from a $\Bad'$ one,
 and vice versa, hence the following result. 
 \alain{trop rapide}
\mathieu{tjr trop rapide ?}

\begin{proposition}\label{srp down up}
{\sc State-resilience} is undecidable for 
%$\post$-ideal-effective ???? and
$\post$-effective,
 $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
when
$\Safe=\mathop{\downarrow} \Safe$.
\end{proposition}

>>>>

Surprisingly, the general undecidability statement regarding resilience had neither been known nor proven.
We show that the resilience problem is undecidable for $\omega^2$-WSTS with strong compatibility. 
%
% \alain{pourquoi n'a-t-on pas prouvé la même chose pour la BR et la k-R ?: où est la difficulté ?}
\begin{proposition}\label{indec WSTS}
{\sc Resilience} for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
\end{proposition}
% \alain{peut-on ajouter des hypothèses d'effectivity ? peut-on montrer que l'exemple n'est pas pred effectif ? l'exemple est-il  ideal-effective (oui pour les propriétes 1 et 2 mais pas 3 car ça permettrait de décider l'arrêt ?)? on voit que l'exemple n'est pas completion-$\post$-effective vu le theoreme qui suit}
\begin{proof}
Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
increasing recursive functions from \cite{FMP-ic04} defined as

@@

%\alain{énoncer pour les reset-VASS c'est plus fort}
\begin{proposition}\label{srp up down}
{\sc State-resilience},
{\sc Bounded-state-resilience} and
{\sc $k$-state-resilience}
are undecidable for reset-VASS, hence for strongly compatible,
$\post$-effective, $\pred$-basis-effective WSTS
when
$\Safe=\mathop{\uparrow} \Safe$.
\end{proposition}

@@

We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
%
Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines.
Indeed, {\sc State-resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
 As seen in 
Proposition~\ref{indec WSTS with dcs},\alain{la Proposition~\ref{indec WSTS with dcs} n'apparait que dans les annéexes...}
 the sequences of $M$ reaching a $\Safe$ state from a $\Bad$ one can be simulated in a reset-VASS
by sequences reaching a $\Safe'$ state from a $\Bad'$ one,
 and vice versa, hence the following result. 
% \alain{trop rapide}
%\mathieu{tjr trop rapide ?}
%\alain{ça va}

\begin{proposition}\label{srp down up}
{\sc State-resilience} is undecidable for reset-VASS, hence also for
%$\post$-ideal-effective ???? and
$\post$-effective,
 $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
when
$\Safe=\mathop{\downarrow} \Safe$.
\end{proposition}
